Distributed Nonlinear Feedback Design with Separable Control-Contraction Metrics

### 2017-03-08 R0V0.tex
Remove cholesky decomposition from notation
Finish example and other minor modifications

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper, large-scale input-affine nonlinear systems are considered. The problem is the design of distributed controllers for this class of network systems. Although control-Lyapunov functions are the natural candidates to design feedback laws, in this case, the construction of these functions is a non-convex optimization problem. Moreover, it may require to verify the positiveness of a function with, at least, $n^2$ elements. In this paper, control-contraction metrics (CCMs) are employed to design controllers with suitable distributed structures. By utilizing a separable structure, the number of verifications needed can be made smaller, as the search of CCMs is formulated as a distributed semidefinite program. An example illustrates the approach for a network system.
\end{abstract}

@@

\mypara Over the years, improvements on distance communication and cheaper computation platforms have allowed the development of large-scale interconnected systems such as smart grids, sensor networks \cite{Pajic2011}, smart manufacturing plants and traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015} and pipe networks \cite{Persis2011a}. One of the main characteristic of these systems is the capacity of being autonomous. Most of current control algorithms are based on the assumption that the system will operate ``near'' to a prescribed set point. Consequently, the system is designed to have fixed behavior. This assumption imposes severe limitations for system, as large excursions are not allowed, in general. For instance, large power variations on smart grids (e.g., when renewable sources of energy are taken into account) are not allowed and manufacturing plants tend to synthesize only a fixed range of goods. To render these systems more flexible, i.e., allowing large excursions a controller algorithm that takes into account a more complex behavior is needed.

\mypara In terms of control theory, the difference between a system operating in a set point and operating in a much larger region lies in the type of equation that describes the system behavior. In the former, the system is described by linear equation while, in the latter, by a nonlinear one. In general, controller design methods for nonlinear systems require the knowledge of a \emph{Lyapunov function}. The search of Lyapunov functions for this class of systems is a non-trivial problem by itself, as no general methodology exists. Another drawback of Lyapunov functions is the fact that they are defined with respect to a particular set point. In case of large excursions, it is more appropriate to define a function depending of the distance between set points. Even if a Lyapunov function is known, no general methodology is known to design distributed feedback laws for nonlinear systems. Thus, existing approaches are, in general, \emph{ad-hoc}. When the design of distributed controllers is considered, the situation is even worse as the problem is known to be, at least, NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}.

@@

\mypara Other works related to analysis under structural constraints to achieve exponential convergence of solutions do exist in the literature. For contraction analysis, see \cite{Aminzare2014a} for analysis according to different network topologies and \cite{Andrieu2016} for transversal contraction analysis. This paper contrast with these works mainly in two points: 1) A  methodology for the design of structured controllers is proposed; 2) Due the formulation of the stabilization problem as a convex optimization problem, an explicit formula of a structured controller is obtained. A particular case of this paper has also been published in \cite{SteinShiromotoManchester2016}. In contrast to this paper, more general conditions to design the distributed controller (cf. Corollary~\ref{cor:main result}) and a methodology to distribute the computation of the feedback law are given.

@@

\paragraph{Notation} Let $\mathbb{S}$ be a totally ordered set and $s_1,s_2,s_3\in\mathbb{S}$. The notation $\mathbb{S}_{[s_1,s_2)}$ (resp. $\mathbb{S}_{\diamond s_3}$) stands for the set $\{s\in\mathbb{S}:s_1\leq s< s_2\}$ (resp. $\{s\in\mathbb{S}:s\diamond s_3\}$, where $\diamond$ is a comparison operator, i.e., $\diamond\in\{<,\geq,=,\ \text{etc}\}$). Let $n>0$ be any integer, the vector $e_i$ denotes the vector with zeros in all rows except the $i$-th where it is 1. A matrix $M\in\mathbb{R}^{n\times n}$ with zero elements except (possibly) those  $m_{ii},\ldots,m_{nn}$ on the diagonal is denoted as $\mathbin{\mathtt{diag}}(m_{ii},\ldots,m_{nn})$. The notation $M\succ 0$ (resp. $M\succeq 0$) stands for $M$ being positive (resp. semi)definite such a class of matrices is denoted as $\mathbf{S}_{\star0}^n=\{M\in\mathbb{R}^{n\times n}:M\star0,M=M^\top\}$, where $\star\in\{\succ,\succeq,\prec,\preceq\}$. For a matrix $M\succ0$, the notation $\mathbin{\mathtt{chol}}(M)$ stands for the \emph{upper-triangular Cholesky factor $U$} satisfying $A=U^\top U$.

@@

\mypara Let $N>0$ be an integer and consider the graph $\mathscr{G}$ defined by the set of vertices $\mathscr{V}=\mathbb{N}_{[1,N]}$ and a set of edges $\mathscr{E}\subset\mathbb{N}_{[1,N]}\times\mathbb{N}_{[1,N]}$. For each index $i\in\mathbb{N}_{[1,N]}$, system~\eqref{eq:general system} can be decomposed into smaller components as follows.
\begin{equation}\label{eq:subsystem}
	\dot{x}_i=f_i(x_i,\breve{x}_i)+b_i(x_i)u_i\;,
\end{equation}
where $\breve{x}_i\in\mathbb{R}^{\breve{n}_i}$ stands for the vector composed of states $x_j$ of systems that are adjacent to system~\eqref{eq:subsystem}. Given a graph $\mathscr{G}_c=(\mathbb{N}_{[1,N]},\mathscr{E}_c)$ that specifies a communication network, for each $i\in\mathscr{V}$. One of the objectives of this work is to design a feedback law $k_i^\ast=k_i(t,x_i,\vec{x}_i,X_i^\ast,\vec{X}_i^\ast,u_i^\ast)$ for \eqref{eq:subsystem} under the constraints defined by $\mathscr{G}_c$.

@@

\mypara Since the vector field $\overline{K}$ is affine on $u$. This implies that, for every fixed $t\geq0$, and for every $S>0$, there exist constants $c_1(S)$ and $c_2(S)$ such that the inequality
	\begin{equation*}
		\left|\overline{K}(t,s,k^\ast)\right|\leq c_1(S)+c_2(S)|k^\ast|
	\end{equation*}
	holds, for every $(s,k^\ast)\in [-S,S]\times\mathbb{R}^m$. Applying \cite[Theorem 2.12]{Teschl2012} pointwisely with respect to $t\geq0$ implies that a solution $k^\ast$ to \eqref{eq:K ODE} is defined for every $s\in[0,1]$.

\mypara Note that, due to the constraint $Y\in\Xi$ and the fact that $W$ is diagonal, each component $i\in\mathbb{N}_{[1,N]}$ of the feedback law $k^\ast$, defined by  Equation~\eqref{eq:contracting feedback law}, is given by
	\begin{equation*}
		k_i^\ast(t,s)=u_i^\ast(t)+\int_0^sK_i(\widetilde{\gamma}_i(t,\sigma),k_i^\ast(t,\sigma))\tfrac{\partial\widetilde{\gamma}_i}{\partial \sigma}(t,\sigma)\,d\sigma\;,
	\end{equation*}
	where $\widetilde{\gamma}_i(\cdot)\in\mathbb{R}^{n_i}\times\mathbb{R}^{\vec{n}_i}$. Thus, $k^\ast$ satisfies the constraint imposed in Problem~\ref{problem formulation}.

@@

@@

\begin{table}
	\begin{tabular}{r|l}
	Method&Time taken by the solver (4 agents)\\ \hline
	Chordal graph decomposition&2.6s (decentralized)\\
	&3.5s (neighbor)\\
	&98s (unconstrained)\\\hline
	\end{tabular}\vspace{1em}
	\caption{A summary of the time taken to solve the matrix inequality~\eqref{eq:MI formulation} using chordal graph decomposition, in the topology illustrated by Fig.~\ref{fig:graph}.}
	\label{tab:chain case}
\end{table}

@@

@@

\mypara The matrix $Y$ have elements given by $Y_{ij}=Y_{ij}(q_i)\delta_{ij}\in\mathbb{R}^{2\times 1}$, for each $i,j\in\mathbb{N}_{[1,4]}$, and for each $q_i\in\mathbb{R}^{2}$, where $\delta_{ij}$ is the Kronecker's delta.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper, large-scale input-affine nonlinear systems are considered. The problem is the design of distributed controllers for this class of network systems. Although control-Lyapunov functions are the natural candidates to design feedback laws, the computation of these functions is a non-convex optimization problem. Moreover, it may require to verify the positiveness of a function with, at least, $n^2$ elements. In this paper, control-contraction metrics (CCMs) are employed to design controllers with prescribed distributed structures. By using a separable structure, the number of verifications needed can be made smaller, as the search of CCMs is formulated as a distributed semidefinite program. An example illustrates the approach for a network system.
\end{abstract}

@@

\mypara Over the years, improvements on distance communication and cheaper computation platforms have allowed the development of large-scale interconnected systems such as smart grids, sensor networks \cite{Pajic2011}, smart manufacturing plants, traffic networks \cite{CanudasdeWitMorbidiLeonOjedaEtAl2015} and pipe networks \cite{Persis2011a}. One of the main characteristic of these systems is the capacity of being autonomous. Most of current control algorithms are based on the assumption that the system will operate ``near'' to a prescribed set point. Consequently, the system is designed to have fixed behavior. This assumption imposes severe limitations for systems, as large excursions are not allowed, in general. For instance, large power variations in smart grids (e.g., when renewable sources of energy are taken into account) are not allowed and manufacturing plants tend to synthesize only a fixed range of goods. To render these systems more flexible, i.e., allowing large excursions a controller algorithm that takes into account the behavior far from the set point is needed.

\mypara In terms of control theory, the difference between a system operating in a set point and operating in a much larger region lies in the type of equation that describes the system behavior. In the former, the system is described by linear equation while, in the latter, by a nonlinear one. In general, controller design methods for nonlinear systems require the knowledge of a \emph{Lyapunov function}. The search of Lyapunov functions for this class of systems is a non-trivial problem by itself, as no general methodology exists. Another drawback of Lyapunov functions is the fact that they are defined with respect to a particular set point. In case of large excursions, it is more appropriate to define a function depending on the distance between points. Even if a Lyapunov function is known, no general methodology is known to design distributed feedback laws for nonlinear systems. Thus, existing approaches are, in general, \emph{ad-hoc}. When the design of distributed controllers is considered, the situation is even worse as the problem is known to be, at least, NP-hard \cite{BlondelTsitsiklis1997,Tanaka2011}.

@@

\mypara Other works related to analysis under structural constraints to achieve exponential convergence of solutions do exist in the literature. For contraction analysis, see \cite{Aminzare2014a} for analysis according to different network topologies and \cite{Andrieu2016} for transversal contraction analysis. This paper contrast with these works mainly in two points: 1) A  methodology for the design of structured controllers is proposed; 2) Due the formulation of the stabilization problem as a convex optimization problem, an explicit algorithm of the distributed controller is provided. A particular case of this paper has also been published in \cite{SteinShiromotoManchester2016}. In contrast to this paper, more general conditions to design the distributed controller (cf. Corollary~\ref{cor:main result}) and a methodology to distribute the computation of the feedback law are given.

@@

\paragraph{Notation} Let $\mathbb{S}$ be a totally ordered set and $s_1,s_2,s_3\in\mathbb{S}$. The notation $\mathbb{S}_{[s_1,s_2)}$ (resp. $\mathbb{S}_{\diamond s_3}$) stands for the set $\{s\in\mathbb{S}:s_1\leq s< s_2\}$ (resp. $\{s\in\mathbb{S}:s\diamond s_3\}$, where $\diamond$ is a comparison operator, i.e., $\diamond\in\{<,\geq,=,\ \text{etc}\}$). Let $n>0$ be any integer, the vector $e_i$ denotes the vector with zeros in all rows except the $i$-th where it is 1. A matrix $M\in\mathbb{R}^{n\times n}$ with zero elements except (possibly) those  $m_{ii},\ldots,m_{nn}$ on the diagonal is denoted as $\mathbin{\mathtt{diag}}(m_{ii},\ldots,m_{nn})$. The notation $M\succ 0$ (resp. $M\succeq 0$) stands for $M$ being positive (resp. semi)definite such a class of matrices is denoted as $\mathbf{S}_{\star0}^n=\{M\in\mathbb{R}^{n\times n}:M\star0,M=M^\top\}$, where $\star\in\{\succ,\succeq,\prec,\preceq\}$.

@@

\mypara Let $N>0$ be an integer and consider the graph $\mathscr{G}$ defined by the set of vertices $\mathscr{V}=\mathbb{N}_{[1,N]}$ and a set of edges $\mathscr{E}\subset\mathbb{N}_{[1,N]}\times\mathbb{N}_{[1,N]}$. For each index $i\in\mathbb{N}_{[1,N]}$, system~\eqref{eq:general system} can be decomposed into smaller components as follows.
\begin{equation}\label{eq:subsystem}
	\dot{x}_i=f_i(x_i,\breve{x}_i)+b_i(x_i)u_i\;,
\end{equation}
where $\breve{x}_i\in\mathbb{R}^{\breve{n}_i}$ stands for the vector composed of states $x_j$ of systems that are adjacent to system~\eqref{eq:subsystem}. Given a graph $\mathscr{G}_c=(\mathbb{N}_{[1,N]},\mathscr{E}_c)$ that specifies a communication network. One of the objectives of this work is to design a feedback law $k_i^\ast=k_i(t,x_i,\vec{x}_i,X_i^\ast,\vec{X}_i^\ast,u_i^\ast)$ for \eqref{eq:subsystem} under the constraints defined by $\mathscr{G}_c$, where $\vec{x}_i\in\mathbb{R}^{\vec{n}_i}$ stands for the vector composed of states $x_j\in\mathbb{R}^{n_j}$, that are adjacent to system $i$, according to the set of edges $\mathscr{E}_c$.

@@

\mypara Since the vector field $\overline{K}$ is affine on $k^\ast$. This implies that, for every fixed $t\geq0$, and for every $S>0$, there exist constants $c_1(S)$ and $c_2(S)$ such that the inequality
	\begin{equation*}
		\left|\overline{K}(t,s,k^\ast)\right|\leq c_1(S)+c_2(S)|k^\ast|
	\end{equation*}
	holds, for every $(s,k^\ast)\in [-S,S]\times\mathbb{R}^m$. Applying \cite[Theorem 2.12]{Teschl2012} pointwisely with respect to $t\geq0$ implies that a solution $k^\ast$ to \eqref{eq:K ODE} is defined for every $s\in[0,1]$.

\mypara Note that, due to the constraint $Y\in\Xi$ and the fact that $W$ is diagonal, each component $i\in\mathbb{N}_{[1,N]}$ of the feedback law $k^\ast$, defined by  Equation~\eqref{eq:contracting feedback law}, is given by
	\begin{equation*}
		k_i^\ast(t,s)=u_i^\ast(t)+\int_0^sK_i(\widetilde{\gamma}_i(t,\sigma))k_i^\ast(t,\sigma)\tfrac{\partial\widetilde{\gamma}_i}{\partial \sigma}(t,\sigma)\,d\sigma\;,
	\end{equation*}
	where $\widetilde{\gamma}_i(\cdot)\in\mathbb{R}^{n_i}\times\mathbb{R}^{\vec{n}_i}$. Thus, $k^\ast$ satisfies the constraint imposed in Problem~\ref{problem formulation}.

@@

\mypara Table~\ref{tab:chain case} shows a summary of the time taken by the solver to compute the matrices $W$ and $Y$ satisfying the set of matrix inequalities~\eqref{eq:main result}, according to the unrestricted, ``neighbor'' and fully decentralized cases.

@@

\begin{table}
	\begin{tabular}{r|c}
	Case&Time taken by the solver (4 agents)\\ \hline
	Unconstrained&98s\\
	Neighbor&3.5s\\
	Fully decentralized&2.6s\\\hline
	\end{tabular}\vspace{1em}
	\caption{A summary of the time taken to solve the set of matrix inequalities~\eqref{eq:main result} using chordal graph decomposition, in the topology illustrated by Fig.~\ref{fig:graph}.}
	\label{tab:chain case}
\end{table}

@@

\mypara The solver took 48s to solve the set of matrix inequalities~\eqref{eq:main result}. The matrix $W$ obtained is given by
\begin{equation*}
	W=\mathbin{\mathtt{diag}}(0.1,1,0.1,1,0.1,1,0.1,1,0.1,1,0.3,1).
\end{equation*}
The matrix $Y$ is given by
\begin{equation*}
	\begin{bmatrix}
		\overline{Y}_{11}(q)&0_{1\times2}&0_{1\times2}&0_{1\times2}&0_{1\times2}&0_{1\times2}\\
		0_{1\times2}&\underline{Y}_{22}(q)&0_{1\times2}&0_{1\times2}&0_{1\times2}&0_{1\times2}\\
		0_{1\times2}&0_{1\times2}&\overline{Y}_{33}(q)&0_{1\times2}&0_{1\times2}&0_{1\times2}\\
		0_{1\times2}&0_{1\times2}&0_{1\times2}&\overline{Y}_{44}(q)&0_{1\times2}&0_{1\times2}\\
		0_{1\times2}&0_{1\times2}&0_{1\times2}&0_{1\times2}&\overline{Y}_{55}(q)&0_{1\times2}\\
		0_{1\times2}&0_{1\times2}&0_{1\times2}&0_{1\times2}&0_{1\times2}&\underline{Y}_{66}(q)
	\end{bmatrix},
\end{equation*}
for every $q\in\mathbb{R}^{12}$, where 
\begin{align*}
 \overline{Y}_{ij}(q)=&\ -\begin{bmatrix}3y_i,&4+4\sum_{i=1}^6(x_i^2+y_i^2)\end{bmatrix},\\
 \underline{Y}_{ij}(q)=&\ -\begin{bmatrix}3y_i,&2+2\sum_{i=1}^6(x_i^2+y_i^2)\end{bmatrix},
\end{align*}
for each index $i,j\in\mathbb{N}_{[1,6]}$.

\mypara In the ``neighbor'' case, the topology of the controller is the same as of the network. More precisely, $\mathscr{G}_c=\mathscr{G}$. The solver took 57s to solve the set of matrix inequalities~\eqref{eq:main result}.

\mypara The matrix $W$ obtained is given by
\begin{equation*}
	W=\mathbin{\mathtt{diag}}(1,1,1,1,1,1,1,1,1,1,2,1).
\end{equation*} 
The matrix $Y$ is given by Equation~\eqref{eq:Y mixed case:neighbor}.

\setcounter{MaxMatrixCols}{20}
\begin{figure*}
\begin{equation}\label{eq:Y mixed case:neighbor}
	\resizebox{.9\textwidth}{!}{
$-\begin{bmatrix}
		2y_1&4+3(|q_1|^2+|\breve{q}_1|^2)&0&0&0&0&0&0&0&0&0&0\\
		0&0&2y_2&2+2(|q_2|^2+|\breve{q}_2|^2)&0&0&0&0&0&0&0&0\\
		0&0&0&0&2y_3&4+3(|q_3|^2+|\breve{q}_3|^2)&0&0&0&0&0&0\\
		0&0&0&0&0&0&2y_4&3+3(|q_4|^2+|\breve{q}_4|^2)&0&0&0&0\\
		0&0&0&0&0&0&0&0&2y_5&4+3(|q_5|^2+|\breve{q}_5|^2)&0&0\\
		0&0&0&0&0&0&0&0&0&0&2y_6&4+3(|q_6|^2+|\breve{q}_6|^2)
	\end{bmatrix}$}
\end{equation}
\end{figure*}

@@

\mypara The matrix $Y$ have elements given by $Y_{ij}=Y_{ij}(q_i)\delta_{ij}\in\mathbb{R}^{2\times 1}$, for each $i,j\in\mathbb{N}_{[1,6]}$, and for each $q_i\in\mathbb{R}^{2}$, where $\delta_{ij}$ is the Kronecker's delta. More specifically, the matrix $Y$ is presented in Equation~\eqref{Y for mixed case}.

\begin{figure*}
\begin{equation}\label{Y for mixed case}
	Y=-\begin{bmatrix}
	2y_1&5+3|q_1|^2&0&0&0&0&0&0\\
 	0&0&2y_2&3+2|q_2|^2&0&0&0&0\\
 	0&0&0&0&2y_3&5+3|q_3|^2&0&0\\
 	0&0&0&0&0&0&2y_4&4+2|q_4|^2
	\end{bmatrix}
\end{equation}
\end{figure*}

\begin{table}
	\begin{tabular}{r|c}
	Case&Time taken by the solver (4 agents)\\ \hline
	Unconstrained&48s\\
	Neighbor&57s\\
	Fully decentralized&43s\\\hline
	\end{tabular}\vspace{1em}
	\caption{A summary of the time taken to solve the set of matrix inequalities~\eqref{eq:main result} using chordal graph decomposition, in the topology illustrated by Fig.~\ref{fig:graph}.}
	\label{tab:mixed example}
\end{table}
